---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# llocalocal

Ten facility-location optimization models (P-Median, P-Center, MCLP,
LSCP, UFCLP, CFLP, DP, UFLP, MAXCAP, PMAXCAP) as integer linear
programs, built as sparse matrices and solved directly via `Rglpk` or
`highs` (the default -- faster than `glpk` on these problems). The
models follow the theoretical framework in `Essai_MarieHelene.pdf`,
the thesis this package originates from.

## Installation

```r
# not on CRAN -- install from the local clone or the private GitHub repo
devtools::install_local(".")
```

## Models

Every model shares the same input contract: `demand`/`candidate`/
`existing_sites` are `sf` POINT layers with an id column, and
`matrix_OD_*` are long origin-destination `data.frame`s
(`from_id`/`to_id`/`distance`) linking them. The examples below all use
the small bundled synthetic dataset (`sample_*`: 15 demand points, 12
candidates, 3 existing sites) so they solve instantly, except MCLP,
which is demonstrated on the real Sherbrooke Bixi dataset bundled with
the package.

### P-Median

Selects exactly `p_facilities` candidate sites to minimize the total
distance between each demand point and its nearest open facility,
weighted by demand (population, jobs, etc.). Sites in `existing_sites`,
if supplied, are forced open ("Required Facilities") instead of
competing for the budget.

```{r p-median}
suppressPackageStartupMessages(library(llocalocal))

res <- p_median(
  demand = sample_demand, demand_id = "id",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  p_facilities = 3
)
res
```

### P-Center

Shares P-Median's setup and constraints, but minimizes the *maximum*
(worst-case) distance between any demand point and its assigned
facility instead of the weighted total -- a minimax rather than a
minisum objective.

```{r p-center}
res <- p_center(
  demand = sample_demand, demand_id = "id",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  p_facilities = 3
)
res
```

### MCLP

Selects up to `p_facilities` candidate sites to maximize the weighted
demand covered within `service_radius`. Unlike LSCP, full coverage
isn't required -- demand outside every open facility's radius is
simply left uncovered. `existing_sites`, if supplied, count as
already-open coverage and don't consume the facility budget.

The package also bundles the real-world case study it originates from:
5,811 candidate sites, 176 demand points, and 25 existing Bixi
bike-share stations in Sherbrooke, QC (see `?bixi-data` for full
provenance and a data-licensing note). `existing_sites` are treated as
Required Facilities -- forced open, and excluded automatically (with a
warning) from `candidate` where the two overlap.

This is a real ~5,800-variable MIP, not a toy example -- it solves in
about 2 seconds with the default `highs` solver.

```{r mclp}
res <- mclp(
  demand = bixi_demand, demand_id = "id", demand_weight = "weight",
  candidate = bixi_candidates, candidate_id = "id",
  existing_sites = bixi_existing, existing_sites_id = "id",
  matrix_OD_candidates = bixi_od_candidates,
  matrix_OD_candidates_from_id = "from_id",
  matrix_OD_candidates_to_id = "to_id",
  matrix_OD_candidates_dist = "travel_time_p50",
  matrix_OD_existing_site = bixi_od_existing,
  matrix_OD_existing_site_from_id = "from_id",
  matrix_OD_existing_site_to_id = "to_id",
  matrix_OD_existing_site_dist = "travel_time_p50",
  cutoff_distance = 130,
  service_radius = 15, p_facilities = 10
)
res
```

10 new stations plus the 25 existing ones cover 67,050 of the 74,200
total weighted demand (population + jobs) -- about 90%, opening only
10 new sites within a 15-minute walk.

### LSCP

The complement of MCLP: minimizes the *number* of facilities needed to
cover every demand point within `service_radius`, with no facility
budget and no partial coverage allowed -- every point must be reached.

```{r lscp}
res <- lscp(
  demand = sample_demand, demand_id = "id",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  service_radius = 6
)
res
```

### UFCLP

Chooses which candidate sites to open, trading off each site's fixed
opening cost (`candidate_fixed_cost`) against the transport cost of
serving demand from it. Unlike P-Median/MCLP, the number of open
facilities isn't fixed -- it falls out of the cost tradeoff, with no
capacity limit on how much demand a site can serve.

```{r ufclp}
res <- ufclp(
  demand = sample_demand, demand_id = "id", demand_weight = "weight",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  candidate_fixed_cost = "fixed_cost"
)
res
```

### CFLP

Extends UFCLP with a capacity limit per facility (`candidate_capacity`):
the total weighted demand assigned to a site can't exceed it. On this
dataset the capacity limit binds -- CFLP has to open more facilities
than UFCLP's unconstrained optimum to fit all the demand in.

```{r cflp}
res <- cflp(
  demand = sample_demand, demand_id = "id", demand_weight = "weight",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  candidate_fixed_cost = "fixed_cost",
  candidate_capacity = "capacity"
)
res
```

### DP

The Dispersion Problem takes no demand layer at all -- it selects
`p_facilities` candidate sites to maximize the *minimum* pairwise
distance between the sites chosen, spreading them out as much as
possible (e.g. siting facilities that shouldn't cluster together).
Since there's no bundled candidate-to-candidate distance table, this
example builds one directly from the candidate geometry with
`sf::st_distance()`.

```{r dp}
cand_dist <- as.matrix(sf::st_distance(sample_candidates))
ids <- sample_candidates$id
od_candidates_dp <- data.frame(
  from_id = rep(ids, times = length(ids)),
  to_id   = rep(ids, each = length(ids)),
  distance = as.numeric(cand_dist)
)
od_candidates_dp <- od_candidates_dp[od_candidates_dp$from_id != od_candidates_dp$to_id, ]

res <- dp(
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = od_candidates_dp,
  p_facilities = 4
)
res
```

### UFLP

The "reverse" of P-Median: the Undesirable Facility Location Problem
selects exactly `p_facilities` sites to *maximize* the total weighted
distance to demand, pushing undesirable installations (landfills,
industrial sites, etc.) as far as possible from where people live.

```{r uflp}
res <- uflp(
  demand = sample_demand, demand_id = "id", demand_weight = "weight",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  p_facilities = 3
)
res
```

### MAXCAP

A competitive-location model: selects up to `p_facilities` new sites to
maximize the weighted demand captured *away from a competitor*
(`existing_sites`), assuming each demand point patronizes whichever
site -- new or competitor's -- is closest to it.

```{r maxcap}
res <- maxcap(
  demand = sample_demand, demand_id = "id", demand_weight = "weight",
  candidate = sample_candidates, candidate_id = "id",
  existing_sites = sample_existing, existing_sites_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  matrix_OD_existing_site = sample_od_existing,
  p_facilities = 3
)
res
```

### PMAXCAP

Extends MAXCAP by making price a decision variable too: the firm
simultaneously chooses where to locate `n_facilities` new sites *and*
what price to charge, to maximize profit against a competitor charging
`competitor_price`. Solved exactly by enumerating the piecewise-linear
profit function's price breakpoints rather than a nonlinear solver.

```{r pmaxcap}
res <- pmaxcap(
  demand = sample_demand, demand_id = "id", demand_weight = "weight",
  candidate = sample_candidates, candidate_id = "id",
  existing_sites = sample_existing, existing_sites_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  matrix_OD_existing_site = sample_od_existing,
  competitor_price = 10, n_facilities = 2
)
res
```
