---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# llocalocal

Ten facility-location optimization models (P-Median, P-Center, MCLP,
LSCP, UFCLP, CFLP, DP, UFLP, MAXCAP, PMAXCAP) as integer linear
programs, solved via `ompr` and `ROI.plugin.glpk`.

## Installation

```r
# not on CRAN -- install from the local clone or the private GitHub repo
devtools::install_local(".")
```

## Example

```{r example}
suppressPackageStartupMessages(library(llocalocal))

res <- p_median(
  demand = sample_demand, demand_id = "id",
  candidate = sample_candidates, candidate_id = "id",
  matrix_OD_candidates = sample_od_candidates,
  p_facilities = 3
)
res
```

## Example with the real Sherbrooke Bixi dataset

The package also bundles the real-world case study it originates from:
5,811 candidate sites, 176 demand points, and 25 existing Bixi
bike-share stations in Sherbrooke, QC (see `?bixi-data` for full
provenance and a data-licensing note). `existing_sites` are treated as
Required Facilities -- forced open, and excluded automatically (with a
warning) from `candidate` where the two overlap.

This is a real ~5,800-variable MIP, not a toy example -- expect it to
take a few minutes to solve with the default `glpk` solver.

```{r bixi-example}
res <- mclp(
  demand = bixi_demand, demand_id = "id", demand_weight = "weight",
  candidate = bixi_candidates, candidate_id = "id",
  existing_sites = bixi_existing, existing_sites_id = "id",
  matrix_OD_candidates = bixi_od_candidates,
  matrix_OD_candidates_from_id = "from_id",
  matrix_OD_candidates_to_id = "to_id",
  matrix_OD_candidates_dist = "travel_time_p50",
  matrix_OD_existing_site = bixi_od_existing,
  matrix_OD_existing_site_from_id = "from_id",
  matrix_OD_existing_site_to_id = "to_id",
  matrix_OD_existing_site_dist = "travel_time_p50",
  cutoff_distance = 130,
  service_radius = 15, p_facilities = 10
)
res
```
```
#> MCLP | 176 demand points | 5786 candidates | radius = 15 | p = 10 | 25 existing (forced) | solver: glpk
#> Warning message:
#> In mclp(...) :
#>   25 id(s) appear in both `candidate` and `existing_sites`; excluding them from `candidate` since they are already open: 1075, 1076, ..., 1099
#>
#> =======================================================
#>   Model         : MCLP
#>   Solver status : success
#> -------------------------------------------------------
#>   Facilities open : 35
#>   Covered demand   : 67050.00
#> =======================================================
#>
```

10 new stations plus the 25 existing ones cover 67,050 of the 74,200
total weighted demand (population + jobs) -- about 90%, opening only
10 new sites within a 15-minute walk.
